The topos-theoretic multiverse

### 2022-09-27 paper-modal-multiverse.tex
Two basic examples

xxx: examples from slides

>>>>

\begin{prop}Let~$({\leq})$ be a transitive almost-full relation.
Then~$({<})$, where~$x < y \equiv (x \leq y \wedge \neg(y \leq x))$,
is well-founded.\end{prop}

\begin{proof}Everywhere, there can be no infinite descending chain, as any
such would also be good.\end{proof}

Unrolling this proof gives a program of type~$(\textsf{Good} \,|\, []) \to
\prod_{x:X} \textsf{Acc}(x)$.

\begin{thm}[Dickson's lemma] If~$X$ and~$Y$ are almost-full,
so is~$X \times Y$.\end{thm}

\begin{proof}
  \begin{enumerate}
    \item It suffices to verify that the \emph{generic infinite
    sequence}~$\gamma = (\alpha,\beta) : \NN \to X \times Y$ is good. Since
    being good can be put as a geometric implication (in fact, a geometric
    formula) and since \textsc{lem} holds \emph{somewhere}, we may assume~\textsc{lem}.

    \item By \textsc{lem} and well-foundedness, there is a minimal value~$\alpha(i_0)$
    among all values of~$\alpha$. \vspace*{-0.2em}
    Similarly, there is a minimal value~$\alpha(i_1)$ among~$(\alpha(n))_{n >
    i_0}$, a minimal value~$\alpha(i_2)$ among~$(\alpha(n))_{n > i_1}$, and so
    on. By \emph{proximal dependent choice}, we can proximally collect these indices
    into a function~$i : \NN \to \NN$; this switches~\textsc{lem} off. \vspace*{-0.2em}

    \item Switching~\textsc{lem} on again, there is a minimal value~$\beta(i(k_0))$
    among all values of~$\beta \circ i$. Hence~$\gamma$ is good in view of
    \begin{align*}
      \alpha(i(k_0)) &\leq \alpha(i(k_0+1)), &
      \beta(i(k_0)) &\leq \beta(i(k_0+1)). \qedhere
    \end{align*}
  \end{enumerate}
\end{proof}
